### 2020-12-05 latex/acrobot.tex
Updated the rotation section, ready to send it to Manfredi.

Using \(\beta\) as our new time variable produces the time-scaled system
\(\hat{\rho}(\beta) = \rho(\tau(\beta))\) with dynamics
\[
    \begin{cases}
        \dot{\hat{\rho}} = \frac{\dot{\rho}}{\dot{\beta}}
        , \\
        \hat{\rho}(0) = \rho_0
        .
    \end{cases}
\]

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\)rad in a clockwise
direction (\ie \(\beta\) going from \(0\) to \(\sign{\rho_0}2\pi\)).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

@@

Note that \(b(\beta,\rho_0)\) is even in \(\rho_0\), which means
\(S(\rho_0) = S(\abs{\rho_0})\).
Recall the assumption that there exist \(\bar{\rho} > 0\) and
\(\epsilon > 0\) where \(S(\rho_0) \geq \epsilon\) for all 
\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
Hence, we have that \(S(\rho_0) \geq \epsilon\)
for every \(\rho_0\) with \(\sqrt{60m^2gl^3} < \abs{\rho} \leq \bar{\rho}\).
Finally, \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\). 
Defining 
\[
    D := [-\bar{\rho},\sqrt{60m^2gl^3}[ \, \cup \, ]\sqrt{60m^2gl^3},\bar{\rho}]
    ,
\]
these facts all together imply that, for
each \(\rho_0 \in D\),
there exists \(I > 0\) small enough that 
\[
    \begin{cases}
        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
        ,\\
        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
        .\\
    \end{cases}
\]
As with oscillations, this \(I\) may not work for all \(\rho_0 \in D\). We deal
with this issue next.

@@

We now examine the case when \(\rho_0 < 0\), where we wish to show that 
\(P_\mathcal{R}(\rho_0) < \rho_0\).
The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}]\), and 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
\[
    R(-2\pi,\rho_0,I) < I^2 r^\prime
    .
\]
Hence,
\begin{align*}
    P_\mathcal{R}(\rho_0)
    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
    &< \rho_0 - IS(\epsilon) + I^2r^\prime
    .
\end{align*}
Choosing \(I_5 \in ]0,I_4]\) so that 
\((I_5)^2r^\prime - I_5 S(\epsilon) \leq 0\) makes
\(P_\mathcal{R}(\rho_0) < \rho_0\) for all 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}[\).

Finally, choosing \(I^\star \in ]0,\min{I_3,I_5}]\) yields that
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
desired.
Looking at ~\eqref{eqn:rhohat-discrete}, this means all solutions of the
discrete-time system will flow away from the origin until it reaches a magnitude
\(\abs{\rho(n)} = \bar{\rho}\).
Hence, the constrained dynamics will eventually escape the set
\(\omega \subset \mathcal{R}\) defined by
\[
    \omega := \left\{(q_u,p_u) \in \Gamma
        \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho} \right\}
    ,
\]
which is displayed in Figure \ref{fig:acrobot-little-omega}.

@@

\subsection{Energy Gain on \(\Omega\)}
Define \(E_\pi\) to be the energy level set associated with \(E(pi,0)\).
We will prove the acrobot gains energy on the set 
\begin{align*}
    \Omega &= \left\{(q_u,p_u) \in \Gamma
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,\\
    &= \mathcal{O} \cup E_\pi \cup \omega
    .
\end{align*}
There exists some control input \(I_1 > 0\) for which the acrobot is gaining
energy on \(\mathcal{O}\). 
Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits flow
towards \(E_\pi\).
Likewise, there exists some \(I_2 > 0\)for which the acrobot is gaining energy
on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
orbit is flowing away from \(E_\pi\).
Choosing \(I^\star \in ]0,\min{I_1,I_2}]\) means the acrobot will gain energy on
both \(\mathcal{O}\) as well as \(\omega\).

We can prove energy gain on \(\Omega\) by showing two properties of the system.
First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
is of measure zero in \(\Gamma\);
second, that \(E_\pi\) is not an invariant set and therefore is not equal to 
\(\Pi^+\).
If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
\(E_\pi\) will generally swing past \(E_\pi\) into \(\omega\).

To show the first condition, we will simply prove that \((\pi,0)\) is an
unstable equilibrium of the constrained dynamics.
Taking the jacobian of ~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\)
yields
\[
    J = \begin{bmatrix}
        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
        3mgl & mgl\bar{q}_aI
    \end{bmatrix}
    ,
\]
which has the characteristic polynomial
\[
    \det\left(\lambda \Id{2} - J\right)
    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
    .
\]
By Descartes' rule of signs, this polynomial has one root with positive real
part. 
Hence, the equilibrium \((\pi,0)\) is unstable.

To show the second condition, we evaluate \(\dot{E}(q_u,p_u)\) at the point
\((0,\pm\sqrt{60m^2gl^3}) \in E_\pi\).
MATLAB returns the value
\[
    \dot{E}(0,\pm\sqrt{60m^2gl^3}) 
    = \mp\frac{g}{5l}
    \sqrt{60m^2gl^3} \sin\left(\bar{q}_a \arctan(I \sqrt{60m^2gl^3})\right)
    .
\]
This is zero only if \(\bar{q}_a \arctan(I\sqrt{60m^2gl^3}) = \frac{\pi}{2}\), a
situation which occurs only when \(\bar{q}_a \geq 1\) at a single value of
\(I\).
Avoiding this value of \(I\) (if it exists) makes \(E_\pi\) a non-invariant set
as desired.

We know that almost all orbits starting in \(\omega\) will escape \(\Omega\).
Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
almost all orbits initialized in \(\mathcal{O}\) will reach \(E_\pi\) and
eventually swing into \(\omega\), rather than stay in \(E_\pi\) and converge to
\((\pi,0)\). 
We conclude that almost all orbits will depart the energy level set associated with
\(E(0,\bar{\rho})\), thereby escaping \(\Omega\) all together.
Hence, the acrobot is gaining energy on \(\Omega\).

>>>>

Using \(\beta\) as our new time variable (via a time reparameterization
\(t = t(\beta)\)) produces the time-scaled system
\(\hat{\rho}(\beta,\rho_0,I) = \rho(t(\beta),\rho_0,I)\) with dynamics
\[
    \begin{cases}
        \diff{\hat{\rho}}{\beta} = \frac{\dot{\rho}}{\dot{\beta}}
        , \\
        \hat{\rho}(0) = \rho_0
        .
    \end{cases}
\]

@@

Suppose we initialize the acrobot at \((\beta,\rho) = (0,\rho_0)\).
One full rotation amounts to \(\beta\) traversing \(2\pi\) rad in a clockwise
direction; that is, when \(\beta\) goes from \(0\) to \(\sign{\rho_0}2\pi\)).
Since \(\dot{\beta}\) does not change sign on \(\mathcal{R}\),
there is a well-defined Poincar\'{e} map describing how the pseudo-radius
\(\rho\) changes each time the orbit \((q(t),p(t))\) hits the \(p_u\) axis, \ie,
every time \(\beta\) changes by \(2\pi\).
We define this Poincar\'{e} map as
\(P_\mathcal{R}(\rho_0) := \hat{\rho}(\sign{\rho_0}2\pi,\rho_0,I)\),
which expands into
\[
    P_\mathcal{R}(\rho_0) = \rho_0 + I \hat{\rho}_1(\sign{\rho_0}2\pi, \rho_0)
    + R(\sign{\rho_0}2\pi,\rho_0,I)
    .
\]

@@

Recall the assumption that there exist \(\bar{\rho} > 0\) and
\(\epsilon > 0\) where \(S(\rho_0) \geq \epsilon\) for all 
\(\rho_0 \in \,]\sqrt{60m^2gl^3},\bar{\rho}]\).
Let us define
\[
    D := \left[-\bar{\rho},\sqrt{60m^2gl^3}\right[ \, 
    \cup \, 
    \left]\sqrt{60m^2gl^3},\bar{\rho}\right]
    .
\]
Since \(b(\beta,\rho_0)\) is even in \(\rho_0\),
\(S(\rho_0) = S(\abs{\rho_0})\);
hence, \(S(\rho_0) \geq \epsilon\) for every \(\rho_0 \in D\).

Finally, recall that \(R(\sign{\rho_0}2\pi,\rho_0,I)\) is \(O(I^2)\), so 
for each fixed \(\rho_0 \in D\)
there exists \(I > 0\) small enough that 
\( IS(\rho_0) + R(\sign{\rho_0}2\pi,\rho_0,I) > 0\).
For this value of \(I\), we have that
\[
    \begin{cases}
        P_\mathcal{R}(\rho_0) > \rho_0 & \rho_0 > 0
        ,\\
        P_\mathcal{R}(\rho_0) < \rho_0 & \rho_0 < 0
        .\\
    \end{cases}
\]
Unfortunately, this \(I\) may not work for all \(\rho_0 \in D\). We deal
with this issue next.

@@

We now examine the case when \(\rho_0 < 0\), where we wish to show that 
\(P_\mathcal{R}(\rho_0) < \rho_0\).
The remainder \(R(\beta,\rho_0,I)\) is bounded above on the compact set 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}]\), and 
Theorem \ref{thm:khalil-perturbation} asserts that there exists some 
\(I_4 > 0\) and \(r^\prime > 0\) so that, for all \(I \in [-I_4,I_4]\), 
\[
    R(-2\pi,\rho_0,I) < I^2 r^\prime
    .
\]
Hence,
\begin{align*}
    P_\mathcal{R}(\rho_0)
    &= \rho_0 - IS(\rho_0) + R(-2\pi,\rho_0,I)
    , \\
    &< \rho_0 - IS(\epsilon) + I^2r^\prime
    .
\end{align*}
Choosing \(I_5 \in ]0,I_4]\) so that 
\((I_5)^2r^\prime - I_5 S(\epsilon) \leq 0\) makes
\(P_\mathcal{R}(\rho_0) < \rho_0\) for all 
\(\rho_0 \in [-\bar{\rho},-\sqrt{60m^2gl^3}[\).

Finally, choosing \(I^\star \in\, ]0,\min\left\{I_3,I_5\right\}]\) yields that
\(\abs{P_\mathcal{R}(\rho_0)} > \abs{\rho_0}\) for all \(\rho_0 \in D\), as
desired.
Looking at the discrete-time system ~\eqref{eqn:rhohat-discrete}, 
a solution of the system will flow away from the origin until it reaches a
magnitude \(\abs{\rho(n)} = \bar{\rho}\).
Hence, the constrained dynamics will eventually escape the set
\(\omega \subset \mathcal{R}\) defined by
\[
    \omega := \left\{(q_u,p_u) \in \Gamma
    \mid E(\pi,0) < E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,
\]
which is displayed in Figure \ref{fig:acrobot-little-omega}.

@@

\subsection{Energy Gain on \(\Omega\)}
Define \(E_\pi\) to be the energy level set associated with \(E(\pi,0)\).
We will prove the acrobot gains energy on the set 
\begin{align*}
    \Omega &= \left\{(q_u,p_u) \in \Gamma
    \mid E(q_u,p_u) < E(0,\bar{\rho}) \right\}
    ,\\
    &= \mathcal{O} \cup E_\pi \cup \omega
    .
\end{align*}
There exists some control input \(I_1 > 0\) for which the acrobot is gaining
energy on \(\mathcal{O}\). 
Hence, there are no closed orbits in \(\mathcal{O}\) and almost all orbits 
hit \(E_\pi\) in finite time.
Likewise, there exists some \(I_2 > 0\) for which the acrobot is gaining energy
on \(\omega\), so there are no closed orbits in \(\omega\) and almost every
orbit is flowing away from \(E_\pi\).
Choosing \(I^\star \in ]0,\min\left\{I_1,I_2\right\}]\) means the acrobot will gain energy on
both \(\mathcal{O}\) as well as \(\omega\).

We can prove energy gain on \(\Omega\) by showing two properties of the system.
First, that the set \(\Pi^+\) of initial conditions converging to  \((\pi,0)\)
is of measure zero in \(\Gamma\);
second, that \(E_\pi\) is not a positively-invariant set.
If both these conditions are true, then orbits flowing from \(\mathcal{O}\) into
\(E_\pi\) will generally swing past \(E_\pi\) into \(\omega\).

To show the first condition, we will simply prove that \((\pi,0)\) is an
unstable equilibrium of the constrained dynamics.
Taking the Jacobian of the constrained dynamics
~\eqref{eqn:acrobot-constrained-dynamics} at \((\pi,0)\) yields
\[
    J = \begin{bmatrix}
        -\frac{6mgl\bar{q}_aI}{5} & \frac{1 - 2m^2gl^3\bar{q}_a^2 I^2}{5ml^2} \\
        3mgl & mgl\bar{q}_aI
    \end{bmatrix}
    ,
\]
which has the characteristic polynomial
\[
    \det\left(\lambda \Id{2} - J\right)
    = \lambda^2 + \frac{mgl\bar{q}_a I}{5} \lambda - 3g
    .
\]
By Descartes' rule of signs, this polynomial has one root with positive real
part. 
The equilibrium \((\pi,0)\) is therefore unstable, so the set \(\Pi^+\) of
initial conditions converging to \((\pi,0)\) is one-dimensional, and hence is
of measure zero in \(\Gamma\).

For the second condition, let us seek a contradiction by 
supposing that \(E_\pi\) is a positively-invariant set. 
Let \(x(t)\) be an orbit of the constrained dynamics initialized at 
\(x(0) \in \mathal{O}\).
Since the constrained dynamics gain energy on \(\mathcal{O}\), there exists a
time \(T > 0\) such that \(x(T) \in E_\pi\).
By the assumed positive invariance of \(E_\pi\), this means 
\(x(t) \in E_\pi\) for all \(t \in [T,\infty[\).
Since \(E_\pi\) is a closed curve containing the equilibrium \((\pi,0)\) and no
other equilibria, \(x(t)\) must converge to \((\pi,0)\).
The initial condition \(x(0) \in \mathcal{O}\) was arbitrary, so this implies
that \(\mathcal{O} \subset \Pi^+\).
But \(\Pi^+\) is a set of measure zero, while \(\mathcal{O}\) has positive
measure. 
This is a contradiction: \(E_\pi\) cannot be positively-invariant.

Non-invariance of \(E_\pi\) and instability of \((\pi,0)\) implies that 
almost all orbits initialized in \(\mathcal{O}\) will reach \(E_\pi\) and
eventually swing into \(\omega\), rather than stay in \(E_\pi\) and converge to
\((\pi,0)\). 
We know that almost all orbits starting in \(\omega\) will reach 
\(E(0,\bar{\rho})\) in finite time;
we conclude that almost all orbits starting in \(\Omega\) will depart the energy
escaping \(\Omega\) all together.
Hence, the acrobot is gaining energy on \(\Omega\).
